\huge{Male-biased sexual selection, but not sexual dichromatism, predicts speciation

### 2019-07-24 Manuscript_Publication.Rmd
Luke edits. Still working on it

@@

Sexual selection results from competition for matings, fertilisations, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008]. This evolutionary force has been proposed to facilitate speciation through the evolution of divergent signals associated with improved mating success [@Lande_1981; @Lande_1982]. Experiments suggest that sexual selection can influence the evolution of a surprisingly diverse set of traits [@Cally_2019], and may therefore strongly influence extant patterns of species diversity. For instance, in antbirds (Thamnophilidae), genera with complex songs and striking dichromatism are more species-rich [@Seddon_2008]. Divergent mating signals in one sex (usually males) co-evolves with divergent preferences for those signals in the opposite sex, leading to behavioural reproductive isolation [@Safran_2013].

Alongside facilitating the evolution of divergent mating signals, sexual selection can increase the extent populations diverge by enlarging the available phenotypic space for mating signals. These predictions emerge as sexual selection may increase genetic diversity as a result of trade-offs between naturally- and sexually-selected traits or trade-offs under *intra*-locus sexual conflict between male and female trait expressions [@Lorch_2003; @Bonduriansky_2011; @Radwan_2016]. Additionally, sexual selection can have both positive and negative consequences for adaptation [*e.g.*, @Kokko_Brooks_2003; @Whitlock_Agrawal_2009; @Holman_Kokko_2013]. For example, sexual selection can elevate average fitness by helping to purge deleterious mutations [@Agrawal_2001; @Siller_2001] and fix beneficial ones [@Whitlock_2000], which should mitigate extinction risk and facilitate adaptation to environmental change. However, sexual selection might also promote extinction by selecting for male traits that have detrimental pleiotropic effects on female traits [@Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013; @Pischedda_2006], or selecting for male phenotypes with improved mating success but which harm population productivity, such as under-investing in parental care [@RN277].

Importantly, the effects of sexual selection on adaptation and speciation may depend on environmental conditions. Both theoretical and empirical studies suggest that under stressful/changing environments, sexual selection may have greater fitness benefits than under benign environments. Stressful environments strengthen the positive correlation between male mating success and female fitness, reducing the burden of negative pleiotropic effects [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017]. Conversely, benign or stable environments ensure consistent selection, preferentially eroding genetic variation at sexually concordant loci (*i.e.* loci where the optimum genotype is equivalent for both sexes). In these stable environments, genetic variation remains disproportionately at sexually antagonistic loci; causing displacement of male and female traits from their optimum. It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008]. Additionally, sexual selection can reinforce local adaptations through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011]. For instance, beak morphology is a trait that arose under natural selection in several taxa and is now a trait under sexual selection [a.k.a. *magic traits*; @Servedio_2011]. In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007]. Through these synergetic effects sexual selection and natural selection (*e.g.* environmental variation over space and time) may lead to increased speciation rates.

Birds are a speciose and well-characterised group of organisms with remarkable and variable sexually-selected traits such as song and colourful plumage, thus are a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003]. A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011]. However, @Kraaijeveld_2011 found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) when using spectrophotometry measures on taxonomic specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores from human observers.

Here, we investigate the association between sexual selection and diversification in birds while extending previous work in multiple ways. Firstly, multiple macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual dimorphism or dichromatism, but sexual selection in birds can involve many other traits, such as song, or the use of ornaments like long tail feathers, bowers, crests and displays [*e.g.*, @Uy_2000; @Pryke_2001]. In fact, it was recently highlighted that song divergence is inversely correlated to sexual dichromatism in passerines, suggesting that there is a trade-off between signalling modalities [@Cooney_2018]. Our study compares the role of sexual dichromatism and other sexual selection proxies in the generation of species. Additionally, we test the relative contribution of environmental variables to the generation of diversity, and how environment and sexual selection could interact to drive speciation processes [@Long_2012; @Connallon_Hall_2016].

To test the link between different measures of sexual selection and diversification processes we use a large (*n* = 5,812) dataset of sexual dichromatism [estimated from illustrated drawings; @Dale_2015], as well as a reduced but still substantial dataset (*n* = 2,465) that gives a composite measure of male-biased sexual selection, capturing variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015]. These datasets allow us to cover the majority of bird genera. Additionally, we use multiple metrics for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$).  \phantomsection

# Materials and methods

We examined the effect of sexual selection on speciation and extinction rate in 97 % of passerines (*n* = 5,812 species; 58 % of all birds). Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions. Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Electronic Supplementary Material (ESM)}.  \phantomsection

@@

#path analysis
best_model <- readRDS('data/path_model.rds')
```
\phantomsection

## Variability across phylogenetic hypotheses and measures of speciation

Estimates of speciation rates varied across phylogenetic trees, especially in the BAMM rates ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95 % highest posterior density (HPD) interval across 100 trees was more than 20 times larger than the 95 % confidence interval calculated for the MCC tree. This contrasts with variation across trees for the other rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), where the 95 % HPD interval across 1,000 trees was near-equivelant to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}). Given the computational requirements of BAMM, the great majority of studies only derive estimates for a single consensus tree; so the high variability in BAMM estimates among trees is notable. Mean measures of speciation across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}). Details of the BAMM results are presented in the supplementary materials.

PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) were included in model selection. However, no interaction terms were significant or present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}). Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation. Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).   
\phantomsection

We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation. Using three response measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$), the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree. Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}). Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 random trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 random trees for $\lambda_{BAMM}$. The HPD (highest posterior density) intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}). For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).  \phantomsection

## Male-biased sexual selection increases speciation

Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation, we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$. This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}). This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a). Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed variable evidence that range size is negatively associated with speciation. Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$. Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}). Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).  \phantomsection

>>>>

#path analysis
best_model <- readRDS('data/path_model.rds')
```

@@

Sexual selection results from competition between same-sex individuals for mates, their gametes, or associated resources, and is typically stronger in males than females [@Fromhage_2016; @Janicke_2016; @Kokko_Jennions2008]. There is long-standing interest in the relationship between sexual selection and the processes of speciation, extinction, and diversification [e.g. @Lande_1981; @Lande_1982]. Experimental work suggests that sexual selection shapes the evolution of many different traits --- not only the obvious ones such as courthship songs or antlers [@Cally_2019] --- while some evolutionary comparative analyses have found that variation in sexual selection between clades predicts phylogenetic diversity. For instance, in antbirds (Thamnophilidae), genera with complex songs and pronounced sexual dichromatism are more species-rich [@Seddon_2008]. ADD ONE MORE EXAMPLE IF POSSIBLE

Several hypotheses predict a positive relationship across clades between the strength of sexual selection and the diversification rate (or the amount of extant phylogenetic diversity). For example, sexual selection is the mechanism underlying reinforcement, a speciation-promoting process in which members of different phylogenetic lineages evolve to avoid unproductive inter-lineage mating/fertilisation (**REFERENCES**). Additionally, sexual selection often operates on traits that can create reproductive isolation when they differ between lineages, such as sexual signals and preferences, sperm-egg interactions, or genitalia (**Safran_2013?, MORE - @Lande_1981; @Lande_1982 ?**). Furthermore, sexual selection is hypothesised to increase diversification by maintaining trait combinations that would be disadvantageous under pure natural selection (**REFS - Bonduriansky_2011 & Radwan_2016? NB I think the journal name might be wrong for Radwan_2016, it's probably JEB**), and by helping to prevent extinction by purging deleterious mutations [@Agrawal_2001; @Siller_2001], fixing beneficial ones [@Whitlock_2000], and accelerating the rate at which populations can adapt to novel environments [@Lorch_2003; @Candolin_2008; @Cally_2019]. Also, sexual selection is sometimes diversifying (as opposed to directional or stabilising), for example when individuals with rare phenotypes have elevated mating success (**REFS? - Holman et al BMC Evol Biol 2009 about Crozier's paradox is one, and likely has some refs in it**): diversifying selection promotes speciaiton by increasing the rate at which isolated populations become genetically distinct.

Conversely, there are multiple ways in which sexual selection is hypothesised to reduce the diversification rate. Firstly, sexual selection can make extinction more likely by favouring traits that improve mating success but reduce population fitness. For example, species with expensive sexual signals may be less resilient to environmental change [@Kokko_Brooks_2003], and sexual selection often creates a 'tragedy of the commons' by selecting for traits that increase the mating success of individuals while reducing population-wide productivity [e.g. infanticide, harassment, and reduced parental care in mate-seeking males; @RN37; @Holman_Kokko_2013]. Sexual selection might also promote extinction by causing maladaptation ('gender load') in female traits that are genetically correlated with sexually-selected male traits [@Pischedda_2006; @Berger_2014; @Bonduriansky_Chenoweth_2009; @Pennell_2013] **Add paper with Mikatake, Hosken and other. Current Biology, about selection on a male trait in beetles constraining female trait evolution even tho the male trait is male-only**. Secondly, sexual selection might decrease the speciation rate by ensuring that individuals with novel trait values have low reproductive success.

Importantly, the relationship between sexual selection and diversification may depend on the environment. Some theoretical and empirical results suggest that sexual selection has a more positive effect on adaptation and population fitness in variable relative to stable environments [@Berger_2014; @Connallon_Hall_2016; @Long_2012; @Martinossi-Allibert_Savković_2017; @Cally_2019]. The hypothesised mechanism is that stable environments lead to the erosion of genetic variation at sexually concordant loci (*i.e.* loci where the fittest allele is the same in both sexes), while sexually antagonistic loci remain comparatively genetically variable, reducing the genetic correlation between male and female fitness. In turn, this means that the most attractive members of the more competitive sex (typically males) produce less fit offspring of the demographically important sex (usually female), such that sexual selection is less likely to favour female-detrimental, extinction-promoting alleles in variable environments. Furthermore, sexual selection might be especially helpful in preventing extinction in variable environments, because (in the words of Charles Darwin), "sexual selection will give its aid to ordinary selection, by assuring to the most vigorous and best adapted males the greatest number of offspring" (DARWIN REF; see also @Agrawal_2001; @Siller_2001; @Boughman_2002; @Maan_2011; ALSO KOKKO + HEUBEL 2008 model about GxE).

<!-- .; causing displacement of male and female traits from their optimum. It follows from these predictions that populations with stronger sexual selection would have fitness benefits, allowing them to adapt to novel environments [*reviewed in* @Candolin_2008]. Additionally, sexual selection can reinforce local adaptations through mate choice on phenotypes that effectively advertise quality in a specific ecological context [@Boughman_2002; @Maan_2011]. For instance, beak morphology is a trait that arose under natural selection in several taxa and is now a trait under sexual selection [a.k.a. *magic traits*; @Servedio_2011]. In Darwin's finches (*Geospiza fortis*) divergent beak morphology is an adaptation to local food availability that has been maintained through assortative mating [@Huber_2007]. Through these synergetic effects sexual selection and natural selection (*e.g.* environmental variation over space and time) may lead to increased speciation rates.   -->

Birds have been a popular focus for macroevolutionary studies of sexual selection and diversification [*e.g.*, @Barraclough_1995; @Seddon_2008; @Huang_2014; @Seddon_2013; @Morrow_Speciation_2003]. Birds are well-suited for this questions as their phylogenetic relationships are comparateively well-known, and they have diverse mating systems and conspicuous sexually-selected traits such as song and colourful plumage. A 2011 meta-analysis, covering 20 studies of birds and other taxa, found a small but significant positive association between sexual selection and speciation, with the average effect size in birds stronger than in mammals but weaker than in insects and fish [@Kraaijeveld_2011]. However, the meta-analysis found large variation in effect sizes between studies, likely because of differences in the metrics used as estimates of speciation and the strength of sexual selection. More recently, @Huang_2014 found no association between sexual dichromatism and speciation (*n* = 918 species) in a study using spectrophotometric measurements of museum specimens [@Armenta_2008] and tip-rate estimates from a molecular-only phylogeny [@Jetz_2012]. Similarly, @Cooney_2017 found no effect of sexual dichromatism on diversification across 1,306 pairs of species, using dichromatism scores from human observers.

Here, we investigate the association between sexual selection and diversification in birds while extending previous work in multiple ways. Firstly, multiple macroevolutionary studies estimate the strength of sexual selection using proxies such as sexual size dimorphism or dichromatism, but sexual selection in birds can involve many other traits, such as song, or the use of ornaments like long tail feathers, bowers, crests and displays [*e.g.*, @Uy_2000; @Pryke_2001]. In fact, it was recently highlighted that song divergence is inversely correlated with sexual dichromatism in passerines, suggesting that a focus on one signalling modality may cause the mismeasurement sexual selection [@Cooney_2018]. Our study compares the role of sexual dichromatism and other sexual selection proxies in the **generation of species (Really? Isn't the composite of speciation and extinction?)** . Additionally, we test the relative effect of some ecological and environmental variables **(I say "ecological" because of range size)** on diversification, and how environment and sexual selection could interact to drive speciation processes [@Long_2012; @Connallon_Hall_2016].

To investigate the link between different measures of sexual selection and diversification, we draw upon a large (*n* = 5,812 species) dataset of sexual dichromatism [estimated from illustrations; @Dale_2015], as well as a smaller but still substantial dataset (*n* = 2,465) that provides an index of male-biased sexual selection, which uses principal components analysis to capture (co)variation in sexual dimorphism, social polygyny and [lack of] paternal care [@Dale_2015]. These datasets allow us to cover the majority of bird genera. Additionally, we use multiple metrics for estimating speciation and extinction rates at the tips of phylogenetic trees, including BAMM [Bayesian Analysis of Macroevolutionary Mixtures; *see*, @Beaulieu_2015; @Rabosky_Extinction_2016; @Moore_2016; @Rabosky_BAMM_flawed_2017], as well as older but reliable tip-rate statistics, such as diversification rate ($\lambda_{DR}$) and node density ($\lambda_{ND}$). We also check the robustness of the results to phylogenetic uncertainty, by considering the posterior distribution of possible phylogenies rather than restricting our analysis to the single most likely phylogeny **(REF?)**.  \phantomsection

## Variability across phylogenetic trees and speciation rate measures

**Is it possible to move this later in the Results? It's maybe a bad idea to start with confusing, worrying results that don't pertain to our main non-null finding regarding the SS index**

We first estimated speciation rates across the phylogeny, using the maximum clade credibility (MCC) tree, as well as 100 or 1000 trees from the posterior distribution of trees [@Jetz_2012]. Estimates of speciation rates varied between phylogenetic trees, especially for rates calcualted by BAMM ($\lambda_{BAMM}$ and $\mu_{BAMM}$), where the 95% highest posterior density (HPD) interval across 100 trees was >20-fold larger than the 95% confidence interval calculated for the MCC tree. Conversely, for other **DIVERSIFICATION??** rate estimates ($\lambda_{DR}$ and $\lambda_{ND}$), the 95% HPD interval across 1,000 trees was near-identical to the 95 % confidence interval calculated for the MCC tree (\href{https://justincally.github.io/SexualSelection_Speciation/#analysis}{Table S9}). Given the computational requirements of BAMM, the great majority of earlier studies have based their estimates on a single consensus tree, and so the inconsistency of the BAMM estimates between different phylogenies with similar likelihoods is notable. Mean measures of speciation rate across 100 trees were positively correlated between measures (DR - BAMM: *r*=`r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`, DR - ND: *r*=`r round(cor(all.trees.join$DR, all.trees.join$ND), digits = 2)`, ND- BAMM: *r*=`r round(cor(all.trees.join$ND, all.trees.join$mean.lambda), digits = 2)`; \href{https://justincally.github.io/SexualSelection_Speciation/#additional_figures_and_tables}{Figure S15}). Given that the calculation of BAMM rates can be affected by the settings of the run and the use of different priors, we compared the estimate of our MCC tree with that of previous published analyses on birds, and found a high correlation (*r*=`r round(cor(Harvey.vs.us$mean.lambda, Harvey.vs.us$Harvey.lambda, use = "pairwise"), digits = 2)`, \href{https://justincally.github.io/SexualSelection_Speciation/#bamm_measures_of_speciation_and_extinction}{Figure S8}). Details of the BAMM results are presented in the supplementary materials.

**I think this part should go after the next 2 sections, as it's a more confusing being about interactions, and we should preserve the order of ideas from the Intro.**

We next fit PGLS models with interaction terms between each measure of sexual dichromatism/selection and the five measures of environmental variability (seasonal temperature variation, log-range size, long-term temperature variation, spatial temperature variation, and NPP) and compared them using AICc model selection. However, no interaction terms were present in the top models ($\delta$ AICc > 4) for any measure of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) or sexual selection (RGB values, spectrophotometry and male-biased sexual selection; $\delta$ AICc > 4; \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_dr_and_nd}{Table S5}, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_models_on_bamm_estimates}{Table S6}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S10}, \href{https://justincally.github.io/SexualSelection_Speciation/#analysis_using_male-biased_measure_of_sexual_selection}{Table S13}). Thus we found no evidence that the effect of sexual selection on speciation is dependent on our measures of environmental variation. Furthermore, we found no evidence that these environmental factors (seasonal temperature variation, long-term temperature variation, spatial temperature variation, and NPP) predict speciation independently from sexual dichromatism/selection (\autoref{model_results}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}).   
\phantomsection

We found no evidence that species with increased sexual dichromatism have higher or lower rates of speciation. Using three different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$, $\lambda_{BAMM}$) as the response variable, the effect of sexual dichromatism was not significant in any PGLS model based on tip-rate estimates and correlation structures from the MCC tree. Specifically, sexual dichromatism showed no association with $\lambda_{DR}$ ($\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}a) or $\lambda_{ND}$ ($\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). Furthermore, speciation rates from BAMM ($\lambda_{BAMM}$) were also unaffected by sexual dichromatism ($\beta$ = `r formatC(MCC.Lambda.top[["coefficients"]][["SDi"]], format = "e", digits = 3)`, p = `r round(anova(MCC.Lambda.top)[["SDi","p-value"]], digits = 3)`; \autoref{model_results}a). PGLS analyses using sexual dichromatism (*n* = 581) measured by spectrophotometry [@Armenta_2008] yielded results concordant with the full dataset; *i.e.* no association between sexual dichromatism and speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}). Our results from models based on the MCC tree are largely corroborated by model estimates from PGLS analyses of the rates and correlation structures from 1,000 random trees (for $\lambda_{DR}$, $\lambda_{ND}$) and 100 random trees for $\lambda_{BAMM}$. The HPD intervals show model estimates are distributed around zero when using complete taxon sampling models and RGB measures of sexual dichromatism (\autoref{model_results}a, \href{https://justincally.github.io/SexualSelection_Speciation/#pgls_model_summary}{Table S8}). For PGLS models using spectrophotometry-based measures of sexual dichromatism, the estimates from the 100 trees in the $\lambda_{DR}$ models are positively skewed (HPD Interval = `r formatC(Armenta.hpd.DR.top$Colour.discriminability, format = "e", digits = 3)`) but normally distributed around zero for $\lambda_{ND}$ and $\lambda_{BAMM}$ (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}).  \phantomsection

## Male-biased sexual selection increases speciation rate

Based on $\lambda_{DR}$ and $\lambda_{ND}$ tip-rate metrics of speciation, we found a negative association between range size and speciation; that is, species with smaller ranges show marginally higher values for $\lambda_{DR}$ and $\lambda_{ND}$. This negative association was small but significant for models using the MCC tree ($\lambda_{DR}$: $\beta$ = `r formatC(MCC.DR.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.DR.top)[["log(range.size.m2)","p-value"]], digits = 3)`; $\lambda_{ND}$: $\beta$ = `r formatC(MCC.ND.top[["coefficients"]][["log(range.size.m2)"]], format = "e", digits = 3)`, p = `r round(anova(MCC.ND.top)[["log(range.size.m2)","p-value"]], digits = 3)`; \autoref{model_results}a, \autoref{Speciation_Figure}). This association was also evident across the estimates from models using the 1,000 trees ($\lambda_{DR}$: HPD Interval = `r formatC(hpd.DR.top[[2]], format = "e", digits = 3)`; $\lambda_{ND}$: HPD Interval = `r formatC(hpd.ND.top[[2]], format = "e", digits = 3)`; \autoref{model_results}a). Subset models with reduced sample size and different measures of sexual selection --- but the same measure of range size --- showed equivocal evidence that range size is negatively associated with speciation. Range size significantly predicted $\lambda_{DR}$ (\autoref{model_results}b) using data subset for male-biased sexual selection (*n* = 2,465) but not $\lambda_{ND}$ or $\lambda_{BAMM}$. Models using data subset for spectrophotometry-based dichromatism (*n* = 581) gave non-significant estimates for the effect of range size on all measures of speciation (\href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Figure S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S11}, \href{https://justincally.github.io/SexualSelection_Speciation/#subsetted_analysis_with_spectrophotometry_data}{Table S12}). Because the range size dataset is the same across the three data subsets we draw our conclusions from the models with the highest power using near-complete taxon sampling (*n* = 5,812).  \phantomsection

We found evidence that the composite measure of male-biased sexual selection, but not the measure of sexual dichromatism, explains variation in the rate of speciation in passerine birds. The absence of a detectable correlation between sexual dichromatism and speciation rate was consistent across different measures of speciation ($\lambda_{DR}$, $\lambda_{ND}$ and $\lambda_{BAMM}$) and both measures of dichromatism (spectral and RGB). These findings substantiate previous studies performed at smaller scales in birds, where sexual dichromatism measures were obtained using spectrophotometry [@Huang_2014] and human observers [@Cooney_2017] and no association was found. The correlation between speciation rate and the index of male-biased sexual selection (which encapsulates variation in sexual size dimorphism, social polygyny, and [lack of] paternal care) was statistically significant for $\lambda_{DR}$, but not for $\lambda_{ND}$ and $\lambda_{BAMM}$, although the estimated direction and magnitude of the correlation was broadly similar across all speciation measures. Interestingly, we found a consistent negative relationship between range size and speciation rate, at least when this rate was quantified using $\lambda_{DR}$ and $\lambda_{ND}$. This suggests that species with smaller range sizes have increased speciation rates. None of the bioclimatic measures of environmental variability that we investigated (*i.e.*, temperature seasonality, long-term temperature variation, and spatial temperature variation) predicted speciation rate.

**Find a paper about Eclectus parrots to add to the paragraph below, maybe. Males and green and females are bright red, and I think females compete for male parental care, so SS is reversed. Another reason why male-biased SS might not be well-correlated with dichromatism**

Although many comparative studies of birds have used sexual dichromatism as a proxy for the strength of sexual selection [*e.g.*, @Barraclough_1995; @Seddon_2013; @Morrow_Speciation_2003; @Owens_1999; @Huang_2014], we found that sexual dichromatism did not predict any measure of speciation rate, while the index of the male-biased sexual selection predicted $\lambda_{DR}$. Given our results, we suggest that sexual dichromatism may not be a robust proxy for sexual selection at least at this broad scale, and that variation in the association between dichromatism and the strength of sexual selection may obscure any relationship with speciation rate, despite the high power of our comparative study. There are several reasons why the use of sexual dichromatism as a proxy for sexual selection is problematic. Firstly, sexual dichromatism can evolve for reasons other than sexual selection, such as when males and females occupy different ecological niches [@Wallace_1889; @Kottler_1980; @Slatkin_1984; @Shine_1989] or experience different selective pressures in contexts other than competition for mates [@Price_2014]. For example, in superb fairy-wrens (*Malurus cyaneus*) female colouration has probably evolved in response to spatial variation in predation pressure, increasing dichromatism [@Medina_2017]. Ecological selection on sexual dichromatism was implicated by our path analysis, which found that sexual dichromatism is positively affected by temperature seasonality (a measure of environmental variation), albeit weakly. Secondly, colour is but one trait and sexual selection may drive the evolution of sex differences in a wide variety of traits used in mate choice or intrasexual competition [@Miles_2018]. For instance, sexual selection may promote investments in other mating signals (*e.g.* song) that trade-off against plumage colour, leading to variable investment in different sexually selected signals across species [@Cooney_2018]. Lastly, not all plumage colouration honestly reflects mate quality. A recent meta-analysis found that converted carotenoids (but *not* carotenoids sourced from the diet) predict mate quality through improvements in parasite resistance and reproductive success [@Weaver_2018]. Given the importance of honest signaling in sexual selection and the variability in colour production mechanisms across birds, sexual dichromatism in many species may be an unreliable measure of mate quality for female birds and sexual selection for researchers.

\newpage\phantomsection
# Materials and methods

We examined the effect of sexual selection on speciation and extinction rate in 97% of passerines (*n* = 5,812 species; 58% of all birds). Specifically, we (*i*) compiled datasets for sexual dichromatism/selection strength and environmental variability, (*ii*) obtained estimates of speciation and extinction rates across passerines, and (*iii*) conducted phylogenetic generalized least-squares (PGLS) regressions. Analyses are documented with reproducible code in the \href{https://justincally.github.io/SexualSelection_Speciation/}{Electronic Supplementary Material (ESM)}.  \phantomsection

@@
